\documentclass[12pt]{article}
\usepackage{array}
\usepackage{geometry}
\geometry{a4paper, margin=1in}
\usepackage{hyperref}
\usepackage{longtable}
\usepackage{amssymb} % For \checkmark
\usepackage{booktabs} % For better table formatting

\title{Applications Plan for Bug Databases}
\author{}
\date{}

\begin{document}

\maketitle

\section*{Database List and Suitability}
\begin{longtable}{|p{3cm}|p{3cm}|p{3cm}|p{3cm}|p{3cm}|}
\hline
\textbf{Database} & \textbf{Bug Classification} & \textbf{Bug Localization} & \textbf{Bug Fixing} & \textbf{Test Case Generation} \\
\hline
\textbf{Bugzilla} & \checkmark & \checkmark & \checkmark & Limited (No test cases, use descriptions to infer) \\
\hline
\textbf{Defects4J} & Not Suitable & Partially Suitable (For localized datasets) & \checkmark & \checkmark (Includes failing/passing test cases) \\
\hline
\textbf{GitHub Issues} & \checkmark & Partially Suitable (Requires linking to code) & Partially Suitable (Limited fixes) & Limited (Test cases depend on repository availability) \\
\hline
\textbf{Apache Bugs} & \checkmark & \checkmark & \checkmark & Limited (Few test cases available) \\
\hline
\end{longtable}

\section*{Applications Plan}

\subsection*{Phase 1: Setup (Week 1--2)}
\begin{itemize}
    \item \textbf{Database Setup:}
    \begin{itemize}
        \item Collect datasets (e.g., Bugzilla API, Defects4J download, GitHub Issues scraping).
        \item Preprocess data (clean text, align buggy/fixed pairs, organize metadata).
    \end{itemize}
    \item \textbf{Environment:}
    \begin{itemize}
        \item Set up modeling frameworks (e.g., Hugging Face Transformers, PyTorch).
        \item Install necessary libraries for handling textual and code data.
    \end{itemize}
\end{itemize}

\subsection*{Phase 2: Application Development (Weeks 3--10)}
\textbf{A. Bug Classification (Weeks 3--4)}
\begin{itemize}
    \item Preprocess:
    \begin{itemize}
        \item Clean textual data from Bugzilla and GitHub Issues.
        \item Map reports to severity or category labels.
    \end{itemize}
    \item Train:
    \begin{itemize}
        \item Fine-tune BERT or CodeBERT for text classification.
    \end{itemize}
    \item Evaluate:
    \begin{itemize}
        \item Use accuracy, F1-score, and confusion matrix.
    \end{itemize}
\end{itemize}

\textbf{B. Bug Localization (Weeks 5--6)}
\begin{itemize}
    \item Preprocess:
    \begin{itemize}
        \item Map bug descriptions to code files/modules (Bugzilla, Apache Bugs).
    \end{itemize}
    \item Train:
    \begin{itemize}
        \item Combine textual data (description) with code embeddings (CodeBERT, Graph Neural Networks).
    \end{itemize}
    \item Evaluate:
    \begin{itemize}
        \item Measure precision and recall for pinpointing buggy locations.
    \end{itemize}
\end{itemize}

\textbf{C. Bug Fixing (Weeks 7--8)}
\begin{itemize}
    \item Preprocess:
    \begin{itemize}
        \item Extract buggy/fixed code pairs (Defects4J).
    \end{itemize}
    \item Train:
    \begin{itemize}
        \item Use CodeT5 or Codex for sequence-to-sequence training.
    \end{itemize}
    \item Evaluate:
    \begin{itemize}
        \item Use BLEU score, patch correctness, and regression testing.
    \end{itemize}
\end{itemize}

\textbf{D. Test Case Generation (Weeks 9--10)}
\begin{itemize}
    \item Preprocess:
    \begin{itemize}
        \item Use test cases in Defects4J to create training pairs.
        \item Extract metadata (e.g., bug description + corresponding test cases).
    \end{itemize}
    \item Train:
    \begin{itemize}
        \item Fine-tune models like Codex to generate test cases.
    \end{itemize}
    \item Evaluate:
    \begin{itemize}
        \item Assess test coverage and fault detection rate.
    \end{itemize}
\end{itemize}

\subsection*{Phase 3: Analysis and Iteration (Weeks 11--12)}
\begin{itemize}
    \item Compare results across datasets for each application.
    \item Identify gaps and iterate on preprocessing or model architectures.
\end{itemize}

\subsection*{Phase 4: Reporting (Weeks 13--14)}
\begin{itemize}
    \item Prepare a research report or paper:
    \begin{itemize}
        \item Include findings, methodologies, and benchmarks.
    \end{itemize}
    \item Explore publication:
    \begin{itemize}
        \item Target software engineering journals or conferences like ICSE, ASE, or EMNLP.
    \end{itemize}
\end{itemize}

\end{document}
